\LARGE \bf Stochastic Uncertainty Propagation in Power System Dynamics using Measure-valued Proximal Recursions

### 2021-06-15 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updated TPS manuscript

@@

The corresponding macroscopic dynamics governing the flow of the joint state PDF $\rho(t,\bm{\theta},\bm{\omega})$ is given by a kinetic Fokker-Planck or Kolmogorov's forward partial differential equation (PDE)
\begin{align}
\frac{\partial \rho}{\partial t} &= - \langle \bm{\omega},\nabla_{\bm{\theta}}\rho \rangle  + \nabla_{\bm{\omega}} \cdot \left( \rho \left( \bm{M}^{-1}\bm{\Gamma}\bm{\omega} + \bm{M}^{-1}\nabla_{\bm{\theta}}V(\bm{\theta}) \right.\right.\nonumber\\
&\qquad\qquad\qquad \left.\left. +\frac{1}{2} \bm{M}^{-1}\bm{\Sigma}\bm{\Sigma}^{\top}\bm{M}^{-1} \nabla_{\bm{\omega}} \log \rho \right) \right),
\label{KineticFPK}	
\end{align}
subject to the initial condition \eqref{FPKInitCond} and the joint parameter PDF $\rho_{\text{param}}$. A direct numerical solution of this PDE initial value problem using conventional discretization (e.g., finite difference) or function approximation techniques will not be scalable in general, as explained in Sec. \ref{subsubsec:technicalchallenges}. In the next Section, we discuss how a measure-valued variational recursion proposed in our recent works \cite{caluya2019ACC,caluya2019TAC,halder2020hopfield,caluya2021TAC} can be employed to address this challenge.

@@

\subsection{From Anisotropic to Isotropic Degenerate Diffusion}\label{subsec:FromAnisoToIso}
Consider the $2n\times 2n$ matrix
\begin{align}
\bm{\Psi}:= \bm{I}_{2} \otimes \bm{M}\bm{\Sigma}^{-1}, \label{DefPsi}	
\end{align}
and define the invertible linear map
\begin{align}
\begin{pmatrix}
\bm{\theta}\\
\bm{\omega}	
\end{pmatrix} \mapsto \begin{pmatrix}
\bm{\xi}\\
\bm{\eta}	
\end{pmatrix} := \bm{\Psi} \begin{pmatrix}
\bm{\theta}\\
\bm{\omega}	
\end{pmatrix}. \label{ChangeOfVar}
\end{align}
Applying It\^{o}'s lemma \cite[Ch. 4.2]{oksendal2013stochastic} to the map \eqref{ChangeOfVar}, and using \eqref{ItoSDEvectorlevel}, we find that the transformed state vector $(\bm{\xi},\bm{\eta})^{\top}$ solves the It\^{o} SDE 
\begin{align}
\begin{pmatrix}
{\mathrm{d}}\bm{\xi}\\
{\mathrm{d}}\bm{\eta}	
\end{pmatrix}
\! = \!\begin{pmatrix}
\bm{\eta}\\
-\nabla_{\bm{\xi}}U(\bm{\xi}) -\nabla_{\bm{\eta}}F\left(\bm{\eta}\right)  	
\end{pmatrix}{\mathrm{d}}t + \!\begin{pmatrix}
 \bm{0}_{n\times n}\\
 \bm{I}_{n}	
 \end{pmatrix}{\mathrm{d}}\bm{w},
\label{XiEtaVectorSDE}	
\end{align}
where the potentials
{\small{\begin{subequations}
\begin{align}
U(\bm{\xi}) &:= \!\displaystyle\sum_{i=1}^{n}\!\frac{1}{\sigma_{i}}P_{i}\xi_{i} + \!\!\!\! \displaystyle\sum_{(i,j)\in\mathcal{E}}\!\!\dfrac{m_{i}}{\sigma_{i}^{2}}k_{ij}\!\left(\!1 \!- \!\cos\!\left(\!\dfrac{\sigma_{i}}{m_{i}}\xi_{i} + \dfrac{\sigma_{j}}{m_{j}}\xi_{j} - \varphi_{ij}\!\!\right)\!\right), \label{defU}\\
F(\bm{\eta}) &:= \dfrac{1}{2}\langle\bm{\eta},\bm{M}^{-1}\bm{\Gamma}\bm{\eta}\rangle. \label{defF}	
\end{align}
\label{NewPotentials}	
\end{subequations}}}

@@

\subsection{Proximal Update}\label{subsec:ProxUpdate}
Looking at \eqref{NewPotentials} and \eqref{NewPDE}, it is natural to consider the energy functional 
\begin{align}
\Phi(\tilde{\rho}) := \displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}}\left(U(\bm{\xi})+F\left(\bm{\eta}\right) + \frac{1}{2}\log\tilde{\rho}\right)\tilde{\rho}\:\differential\bm{\xi}\differential\bm{\eta},
\label{LyapButNotProx}	
\end{align}
which is the sum of a potential energy (expected value of $U$), a kinetic energy (expected value of $F$), and an internal energy (scaled negative entropy $-\int\tilde{\rho}\log\tilde{\rho}$). Indeed, the functional \eqref{LyapButNotProx} is a Lyapunov functional along the solution of \eqref{NewPDE}. To see this, ...

>>>>

Our technical approach juxtaposes with respect to the existing literature discussed above, in that we approximate neither the dynamical nonlinearity nor the statistics. Instead of treating the exact nonlinearity as bane, we exploit the geometry induced by the power system's dynamical nonlinearity

@@

The corresponding macroscopic dynamics governing the flow of the joint state PDF $\rho(t,\bm{\theta},\bm{\omega})$ is given by a kinetic Fokker-Planck partial differential equation (PDE)
\begin{align}
\frac{\partial \rho}{\partial t} &= - \langle \bm{\omega},\nabla_{\bm{\theta}}\rho \rangle  + \nabla_{\bm{\omega}} \cdot \left( \rho \left( \bm{M}^{-1}\bm{\Gamma}\bm{\omega} + \bm{M}^{-1}\nabla_{\bm{\theta}}V(\bm{\theta}) \right.\right.\nonumber\\
&\qquad\qquad\qquad \left.\left. +\frac{1}{2} \bm{M}^{-1}\bm{\Sigma}\bm{\Sigma}^{\top}\bm{M}^{-1} \nabla_{\bm{\omega}} \log \rho \right) \right),
\label{KineticFPK}	
\end{align}
subject to the initial condition \eqref{FPKInitCond} and the joint parameter PDF $\rho_{\text{param}}$. A direct numerical solution of this PDE initial value problem using conventional discretization (e.g., finite difference) or function approximation techniques will not be scalable in general, as explained in Sec. \ref{subsubsec:technicalchallenges}. In the next Section, we discuss how a measure-valued variational recursion proposed in our recent works \cite{caluya2019ACC,caluya2019TAC,halder2020hopfield,caluya2021TAC} can be employed to address this challenge.

@@

\subsection{From Anisotropic to Isotropic Degenerate Diffusion}\label{subsec:FromAnisoToIso}
Consider the $2n\times 2n$ matrix
\begin{align}
\bm{\Psi}:= \bm{I}_{2} \otimes \left(\bm{M}\bm{\Sigma}^{-1}\right), \label{DefPsi}	
\end{align}
and define the invertible linear map
\begin{align}
\begin{pmatrix}
\bm{\theta}\\
\bm{\omega}	
\end{pmatrix} \mapsto \begin{pmatrix}
\bm{\xi}\\
\bm{\eta}	
\end{pmatrix} := \bm{\Psi} \begin{pmatrix}
\bm{\theta}\\
\bm{\omega}	
\end{pmatrix}. \label{ChangeOfVar}
\end{align}
Applying It\^{o}'s lemma \cite[Ch. 4.2]{oksendal2013stochastic} to the map \eqref{ChangeOfVar}, and using \eqref{ItoSDEvectorlevel}, we find that the transformed state vector $(\bm{\xi},\bm{\eta})^{\top}$ solves the It\^{o} SDE 
\begin{align}
\begin{pmatrix}
{\mathrm{d}}\bm{\xi}\\
{\mathrm{d}}\bm{\eta}	
\end{pmatrix}
\! = \!\begin{pmatrix}
\bm{\eta}\\
-\nabla_{\bm{\xi}}U(\bm{\xi}) -\nabla_{\bm{\eta}}F\left(\bm{\eta}\right)  	
\end{pmatrix}{\mathrm{d}}t + \!\begin{pmatrix}
 \bm{0}_{n\times n}\\
 \bm{I}_{n}	
 \end{pmatrix}{\mathrm{d}}\bm{w},
\label{XiEtaVectorSDE}	
\end{align}
where the potentials
{\small{\begin{subequations}
\begin{align}
U(\bm{\xi}) &:= \!\displaystyle\sum_{i=1}^{n}\!\frac{1}{\sigma_{i}}P_{i}\xi_{i} + \!\!\!\! \displaystyle\sum_{(i,j)\in\mathcal{E}}\!\!\dfrac{m_{i}}{\sigma_{i}^{2}}k_{ij}\!\left(\!1 \!- \!\cos\!\left(\!\dfrac{\sigma_{i}}{m_{i}}\xi_{i} + \dfrac{\sigma_{j}}{m_{j}}\xi_{j} - \varphi_{ij}\!\!\right)\!\right), \label{defU}\\
F(\bm{\eta}) &:= \dfrac{1}{2}\langle\bm{\eta},\bm{M}^{-1}\bm{\Gamma}\bm{\eta}\rangle. \label{defF}	
\end{align}
\label{NewPotentials}	
\end{subequations}}}

@@

\subsection{Proximal Update}\label{subsec:ProxUpdate}
Looking at \eqref{NewPotentials} and \eqref{NewPDE}, it is natural to consider the energy functional 
\begin{align}
\Phi(\tilde{\rho}) := \displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}}\left(U(\bm{\xi})+F\left(\bm{\eta}\right) + \frac{1}{2}\log\tilde{\rho}\right)\tilde{\rho}\:\differential\bm{\xi}\differential\bm{\eta},
\label{LyapButNotProx}	
\end{align}
which is the sum of a potential energy (expected value of $U$), a kinetic energy (expected value of $F$), and an internal energy (scaled negative entropy, the entropy being $-\int\tilde{\rho}\log\tilde{\rho}$). Indeed, the functional \eqref{LyapButNotProx} is a Lyapunov functional along the solution of \eqref{NewPDE}. To see this, ...
